---
title: "expedia"
output: html_document
---


```{r, echo=FALSE, message=FALSE}
library(ggplot2)    # Data visualization
library(data.table) # Faster data reading
library(dplyr)      # Data aggregation etc.
library(scales)     # Plot scaling
library(corrplot)   # Correlations
```

# Loading and Transforming Data

```{r, echo=FALSE}
train <- fread("xzcat ../data/train1M.csv.xz", header = TRUE)

# convert to factors
cols <- colnames(train)
nonfactor <- c("date_time", "srch_ci", "srch_co", "orig_destination_distance")
for (i in cols[!cols %in% nonfactor] ) {
  train[[i]] <- as.factor(train[[i]])
}

# convert to dates
train$date_time <- as.POSIXct(train$date_time)
train$srch_ci <- as.Date(train$srch_ci)
train$srch_co <- as.Date(train$srch_co)
train$date <- as.Date(train$date_time)
train$Year  <- as.numeric(format(train$date, "%Y"))
train$Month <- as.numeric(format(train$date, "%m"))
```

## Descriptive Statistics

```{r, echo=FALSE}
str(train)
```

```{r, echo=FALSE}

grouping <- function(cat, cat_name) {
  train.agg <- train %>% group_by(Year, Month, cat) %>% summarize(count=n())
  train.agg$date <- as.POSIXct(paste(train.agg$Year, train.agg$Month, "01", sep = "-")) # set date to first of month
  
  ggplot(train.agg, aes(x = date, y = count, group = cat, col = cat)) +
  geom_line(size=0.2) +
#  facet_wrap(~Category, ncol = 5, scales = "free_y") +
  theme(legend.title=element_text(cat_name),
        axis.text=element_text(size=7),
        strip.text = element_text(size = 9)) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(title=paste("Daily events:", cat_name))
}

grouping(train$is_booking, "is_booking")
grouping(train$is_mobile, "is_mobile")
grouping(train$is_package, "is_package")
grouping(train$channel, "channel")
grouping(train$srch_adults_cnt, "srch_adults_cnt")
grouping(train$srch_children_cnt, "srch_children_cnt")

grouping(train$posa_continent, "posa_continent")


#table(train$hotel_continent)
#table(train$hotel_country)
```

* Exploration
  * site/country
  * channel
  * continent
  * country/region/city
  * user bookings
  * distance
  * mobile
  * package
  * adults/children


* Engineering
  * Day of Week
  * Morning Evening
  * duration of stay
  * time to departure
  * domestic/ continental / intercontinental
  
* customer-specific
  * #bookings
  * #bookings in last year
* hotel-specific


* recommendation
  *