---
title: "Kaggle Expedia"
output: html_document
---

## [Logs of customer behavior published by Expedia](https://www.kaggle.com/c/expedia-hotel-recommendations)

> These include what customers searched for, how they interacted with search results (click/book), whether or not the search result was a travel package.
> Expedia is interested in predicting which hotel group a user is going to book. Expedia has in-house algorithms to form hotel clusters, where similar hotels for a search (based on historical price, customer star ratings, geographical locations relative to city center, etc) are grouped together.


### Training/Test Data

Column name               | Description                                                                           | Data type
------------------------- | ------------------------------------------------------------------------------------- | ---------
date_time                 | Timestamp                                                                             | string
site_name                 | ID of the Expedia point of sale (i.e. Expedia.com, Expedia.co.uk, Expedia.co.jp, ...) | int
posa_continent            | ID of continent associated with site_name                                             | int
user_location_country     | The ID of the country the customer is located                                         | int
user_location_region      | The ID of the region the customer is located                                          | int
user_location_city        | The ID of the city the customer is located                                            | int
orig_destination_distance | Physical distance between a hotel and a customer at the time of search. \
                            A null means the distance could not be calculated                                     | double
user_id                   | ID of user                                                                            | int
is_mobile                 | 1 when a user connected from a mobile device, 0 otherwise                             | tinyint
is_package                | 1 if the click/booking was generated as a part of a package \
                           (i.e. combined with a flight), 0 otherwise                                             | int
channel                   | ID of a marketing channel                                                             | int
srch_ci                   | Checkin date                                                                          | string
srch_co                   | Checkout date                                                                         | string
srch_adults_cnt           | The number of adults specified in the hotel room                                      | int
srch_children_cnt         | The number of (extra occupancy) children specified in the hotel room                  | int
srch_rm_cnt               | The number of hotel rooms specified in the search                                     | int
srch_destination_id       | ID of the destination where the hotel search was performed                            | int
srch_destination_type_id  | Type of destination                                                                   | int
hotel_continent           | Hotel continent                                                                       | int
hotel_country             | Hotel country                                                                         | int
hotel_market              | Hotel market                                                                          | int
is_booking                | 1 if a booking, 0 if a click                                                          | tinyint
cnt                       | Numer of similar events in the context of the same user session                       | bigint
hotel_cluster             | ID of a hotel cluster                                                                 | int

### Destinations

Column name         | Description                                                | Data type
------------------- | ---------------------------------------------------------- | ---------
srch_destination_id | ID of the destination where the hotel search was performed | int
d1-d149             | latent description of search regions                       | double


# Loading and Transforming Data

```{r, echo=FALSE, message=FALSE}
library(ggplot2)    # Data visualization
library(data.table) # Faster data reading
library(dplyr)      # Data aggregation etc.
library(scales)     # Plot scaling
library(corrplot)   # Correlations
```

```{r, echo=FALSE, message=FALSE}
train <- fread("xzcat ../data/train1M.csv.xz", header = TRUE)

# convert to factors
cols <- colnames(train)
nonfactor <- c("date_time", "srch_ci", "srch_co", "orig_destination_distance")
for (i in cols[!cols %in% nonfactor] ) {
  train[[i]] <- as.factor(train[[i]])
}

# convert to dates
train$date_time <- as.POSIXct(train$date_time)
train$srch_ci <- as.Date(train$srch_ci)
train$srch_co <- as.Date(train$srch_co)
train$date <- as.Date(train$date_time)
#train$week <- cut(train$date, "weeks")
train$Year  <- as.numeric(format(train$date, "%Y"))
train$Month <- as.numeric(format(train$date, "%m"))

# more feature engineering
train$durStay <- as.numeric(train$srch_co - train$srch_ci)
train$durStay[train$srch_ci > train$srch_co] <- -1

train$timeToDep <- as.numeric(train$srch_ci - train$date)
train$timeToDep[train$timeToDep < 0] <- -1
train$timeToDep[train$timeToDep > 1000] <- -1
```

## Descriptive Statistics

```{r, echo=FALSE}
str(train)
```

```{r, echo=FALSE}

datePlot <- function(cat_name) {
  column <- train[[cat_name]]
  train.agg <- train %>% group_by(date, column) %>% summarize(count=n())
#  train.agg <- train %>% group_by(Year, Month, cat) %>% summarize(count=n())
#  train.agg$date <- as.POSIXct(paste(train.agg$Year, train.agg$Month, "01", sep = "-")) # set date to first of month
  
  ggplot(train.agg, aes(x = date, y = count, group = column, color = column)) +
    geom_line(size=0.2) +
    theme(axis.text=element_text(size=7),
          strip.text = element_text(size = 9)) +
    scale_y_continuous(limits = c(0, NA)) +
    scale_color_discrete(name = cat_name) + # name of line legend
    labs(title=paste("Daily events:", cat_name))
}

barPlot <- function(cat_name, legendpos = "right") {
  column <- train[[cat_name]]
  ggplot(train, aes(x=column, fill=column)) + geom_bar() +
    labs(title=paste("Number of", cat_name)) +
    theme(legend.position=legendpos)
}

datePlot("is_booking")
datePlot("is_mobile")
datePlot("is_package")
datePlot("channel")
datePlot("srch_adults_cnt")
datePlot("srch_children_cnt")
datePlot("posa_continent")

barPlot("posa_continent")
barPlot("is_booking")
barPlot("is_mobile")
barPlot("is_package")
barPlot("channel")
barPlot("srch_adults_cnt")
barPlot("srch_children_cnt")
barPlot("srch_rm_cnt")
barPlot("cnt")
barPlot("hotel_cluster", "none")
barPlot("hotel_continent")
barPlot("hotel_country", "none")
barPlot("hotel_market", "none")

barPlot("durStay")

dur <- train %>% group_by(durStay) %>% summarize(count=n())
ggplot(dur, aes(x=durStay, y=count)) + geom_line()

ttd <- train %>% group_by(timeToDep) %>% summarize(count=n())
ggplot(ttd, aes(x=timeToDep, y=count)) + geom_line()

users <- data.frame(table(train[train$is_booking == 1]$user_id))
colnames(users) <- c("user_id", "freq")
usr <- users %>% group_by(freq) %>% summarize(count=n())
ggplot(usr, aes(x=freq, y=count)) + geom_line()
```

* Exploration
  * site/country
  * channel
  * continent
  * country/region/city
  * user bookings
  * distance
  * mobile
  * package
  * adults/children


* Engineering
  * Day of Week
  * Morning Evening
  * duration of stay
  * time to departure
  * domestic/ continental / intercontinental
  
* customer-specific
  * #bookings
  * #bookings in last year
* hotel-specific


* recommendation
  *